The Evolution of Phenotypically Invariant Genetic Networks

### 2017-07-28 handnotes.tex
started a metabolic example

>>>>

  \subsection*{Metabolic network}

  Consider an organism that can metabolize two different sugars $s_{1}$ and $s_{2}$ (present at logarithmic concentrations $u_{1}$ and $u_{2}$ respectively), with enzymes $e_{1}$ and $e_{2}$ (with log concentration denoted as $\phi_{1}$ and $\phi_{2}$). Further suppose that one of the sugars $s_{2}$ is the preferred energy source (perhaps it contains significantly more energy than the other sugar or is otherwise more efficiently metabolized). The organism may have a gene regulatory network $\mathcal{S}$ that can depoly a situation specific metabolic strategy. That is depending on both $u{2}$ and $u_{1}$ the organism synthesize an appropriate $\phi_{1}$ and $\phi_{2}$. Furthermore, consider this system to contain at least two transcription factors, whose log concentrations are given by $\kappa_{1}$, $\kappa_{2}$, $\dots \kappa_{n}$.

  Minimally such a system may have the architecture,
  \begin{align*}
    \mathcal{S} = \left\{ \begin{array}{ll} \dot{\kappa}(t) &= \begin{bmatrix} 0 & -1 \\ 0 & 0 \end{bmatrix} \begin{bmatrix} \kappa_{1} \\ \kappa_{2} \end{bmatrix} (t) + \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix} \begin{bmatrix} u_{1} \\ u_{2} \end{bmatrix} (t) \\[11pt]
      \phi(t) &= \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix} \begin{bmatrix} \kappa_{1} \\ \kappa_{2} \end{bmatrix} (t) \end{array} \right.
  \end{align*}
  $\mathcal{S}$ is minimal as its reachability and observability matrices both have rank $=2$, $\text{rank}(\mathcal{R}) = \text{rank}(\mathcal{O}) = 2 = n \implies \min$.

  The impulse response matrix of this system is, 
  \begin{align*}
    h(t) = \begin{bmatrix} 1 & - t \\ 0 & 1 \end{bmatrix}  .
  \end{align*}

  The phenotype is, 
  \begin{align*}
    \phi(t) = \exp \left\{ \kappa(0) + \int_{0}^{t} h(t) u(t) dt \right\} 
  \end{align*}
  ($\kappa(0)$ can be set to something like $\left[ -10, -10 \right]$, assuming the transient transcription factor and enzyme concentrations in the organism are typically quite low).

  We can see that changing coordinates on this system will find new system architecutres, as shown before, however, we can also apply the Kalman decomposition to find systems in different dimensions -- that is systems that employ more than $2$ transcription factors, yet have the same external dynamics/phenotype.

  DO KALMAN EXAMPLE HERE.
